---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cmvdrg

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/ucl-pathgenomics/cmvdrg.svg?branch=master)](https://travis-ci.org/ucl-pathgenomics/cmvdrg)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.0-6666ff.svg)](https://cran.r-project.org/)
![GitHub repo size](https://img.shields.io/github/repo-size/ucl-pathgenomics/cmvdrg.svg)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/ucl-pathgenomics/cmvdrg.svg)
<!-- badges: end -->



## overview

cmvdrg is a R package to enable antiviral drug resistance genotyping, with Human Cytomegalovirus sequencing data. Accepted inputs are FASTA whole genomes, FASTA fragments which will be mapped to Refseq NC_006273.2. NGS variant data assembled to NC_006273.2 is accepted in VCF >= ver4.0 & Varscan2 tab formats. 


#### Database
Contains the relationships between:

 - Mutation
 - Drug susceptibility, values are EC50 fold changes to susceptible strains. "Susceptible" or "Resistant" are present if data is anecdotal.
 - The method for Resistance Phenotyping, including where possible strian information used in marker transfer. i.e. AD169, TOWNE.
 - Publication reference, DOI, web address & review paper reference if used.
 - Where a co-mutation susceptibility profile is available for a known drug, this is included.
 - Administrative information, last update to reference, updater, whether the row is used in calculations or not. i.e. if a mutations is present in a review paper, but is not traceable to given references this column is set to "U" Unused, or "R" to Review.


#### Web service
A user friendly Shiny Applications has been bundled with this package. The same application is available over the internet here [http://51.11.13.133:3838/cmvdrg/](http://51.11.13.133:3838/cmvdrg/) where the terms of use are contained.


## Installation

You can install the current version from [GitHub](https://github.com/ucl-pathgenomics/cmvdrg) with:

``` r
# install.packages("devtools")
devtools::install_github("ucl-pathgenomics/cmvdrg")
```

Dependendencies for FASTA file handling are MAFFT and SNP-Sites available preffereably via conda. snp-sites >= 2.3 has been tested.
```{bash, eval=FALSE}
conda config --add channels bioconda
conda install snp-sites
conda install mafft
```


## Usage

```{r setup, message = F}
library("cmvdrg")

## call resistance mutations only
my_sample = system.file("testdata", "A10.vcf", package = "cmvdrg")


# this calls any variants in the sample file, that are also present in the cmvdrg database.
data = call_resistance(infile = my_sample, all_mutations = F)


print(data[,c("change", "freq", "Ganciclovir")])



## call all mutations - i.e. there may be interesting non-synonymous mutants predsent in resistance genes.
# returns all variants, with any resistance data, with annotation of protein coded for & if it coded for a change in protein sequence.
mutations_all = call_resistance(infile = my_sample, all_mutations = T)


# to view all mutations in resistance genes we can use base R to filter.
mutations_res = mutations_all[mutations_all$GENEID %in% c("UL54", "UL97", "UL27", "UL51", "UL56", "UL89"),]


# are there any non-synonymous (DNA variants that result in a change of amino acid) variants in resistance genes.
mutations_res_nonsyn = mutations_res[mutations_res$CONSEQUENCE == "nonsynonymous",]

# here the top 3 mutations are nonsynonymous, but have no identified resistance effect.
head(mutations_res_nonsyn[,c(1,8,21,32:40)])


## run the shiny application
# runShinyCMV()

```

## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible example on the GitHub Issues page. For questions and other discussions feel free to contact. [Oscar Charles - maintainer](mailto:oscar.charles.18@ucl.ac.uk)

